---
title: "NLA_MACROS"
author: "David"
date: "11/7/2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

Load packages
```{r}
library(MuMIn)
library(lme4)

#install.packages('reshape')

library(reshape)
library(tidyr)
library(dplyr)

```

Load NLA 2007 data
```{r}

# load data
# note: update working directory as needed.

NLA_2007=read.csv("/Users/davidmanning/Desktop/EmergenceAI/EPA_NARS_LAKES_MACROS_2007.csv")

NLA_2007_landuse=read.csv("/Users/davidmanning/Desktop/EmergenceAI/NLA_2007_landuse.csv")
NLA_2007_waterchem=read.csv("/Users/davidmanning/Desktop/EmergenceAI/NLA_2007_waterchem.csv")
head(NLA_2007_waterchem)
NLA_2007_siteinfo=read.csv("/Users/davidmanning/Desktop/EmergenceAI/NLA_2007_siteinfo.csv")
head(NLA_2007_siteinfo)
NLA_2007_waterquality=read.csv("/Users/davidmanning/Desktop/EmergenceAI/NLA2007_WaterQuality_20091123.csv")

NLA_2007_metrics=read.csv("/Users/davidmanning/Documents/EPA_NLA_Benthic_metrics.csv")
names(NLA_2007_metrics)

dat1=inner_join(NLA_2007,NLA_2007_siteinfo,by="SITE_ID")
dat2=inner_join(dat1,NLA_2007_landuse,"SITE_ID")
dat3=inner_join(dat2,NLA_2007_waterquality,by="SITE_ID")
dat4=inner_join(dat3,NLA_2007_metrics,by="SITE_ID")

```

Explore NLA data
```{r}

head(dat3)
head(dat4)
names(dat4)
str(dat4)
names(dat4)
ggplot() + geom_point(data=dat4,aes(x=YEAR.x,y=NTL))

```

Reshape data with columns by order
```{r}

head(dat3)
names(dat3)
library(reshape)

l.reshape.frame=cast(dat3,SITE_ID+LON_DD+LAT_DD+STATE+LAKEAREA+LAKENAME+HUC_8+BASINAREA_KM2+PCT_DEVELOPED_BSN+PCT_DECID_BSN+PCT_CONIF_BSN+PCT_MIXED_BSN+PCT_FOREST_BSN+PCT_SHRUBLAND_BSN+PCT_GRASS_BSN+PCT_PASTURE_BSN+PCT_CROPS_BSN+PCT_AGRIC_BSN+PCT_WETLAND_BSN+SECMEAN+PH_LAB+COND+ANC+TURB+NH4N_PPM+NO3_NO2+NTL+PTL+CL_PPM+NO3N_PPM+SO4_PPM+CA_PPM+MG_PPM+K_PPM+SIO2+CHLA~FAMILY,value="TOTAL_300",mean)
l.reshape.frame[1:5,]
l.frame1=as.data.frame(l.reshape.frame)
l.frame1[is.na(l.frame1)]=0
ncol(l.frame1)
names(l.frame1)
l.fam.abund=as.data.frame(l.frame1[38:187])
l.fam.abund[1:5,]
l.tot.abund=rowSums(l.fam.abund)
l.tot.abund
l.sites=as.data.frame(l.frame1[1:36])
head(l.sites)

length(l.tot.abund) # n = 1210 sites




```

Classify as emergent/non-emergent macroinvertebrates by family.
```{r}

colnames(l.fam.abund) # list of families
emerge_lake=with(l.fam.abund,(cbind(AESHNIDAE,AMELETIDAE,BAETIDAE,BAETISCIDAE,BRACHYCENTRIDAE,CAENIDAE,CALOPTERYGIDAE,CAPNIIDAE,CERATOPOGONIDAE,CHAOBORIDAE,CHIRONOMIDAE,CHLOROPERLIDAE,COENAGRIONIDAE,CORDULIIDAE,CORYDALIDAE,CRAMBIDAE,CULICIDAE,DIPSEUDOPSIDAE,DOLICHOPODIDAE,EMPIDIDAE,EPHEMERELLIDAE,EPHEMERIDAE,EPHYDRIDAE,GOMPHIDAE,HELICOPSYCHIDAE,HEPTAGENIIDAE,HYDROPSYCHIDAE,HYDROPTILIDAE,LEPIDOSTOMATIDAE,LEPTOCERIDAE,LEPTOHYPHIDAE,LEPTOPHLEBIIDAE,LESTIDAE,LIBELLULIDAE,LIMNEPHILIDAE,MACROMIIDAE,METRETOPODIDAE,MOLANNIDAE,MUSCIDAE,NEMOURIDAE,PERLIDAE,PERLODIDAE,PHILOPOTAMIDAE,PHRYGANEIDAE,POLYCENTROPODIDAE,POLYMITARCYIDAE,PSYCHODIDAE,PSYCHOMYIIDAE,PYRALIDAE,RHYACOPHILIDAE,SCIOMYZIDAE,SERICOSTOMATIDAE,SIALIDAE,SIMULIIDAE,SIPHLONURIDAE,SISYRIDAE,STRATIOMYIDAE,TABANIDAE,TAENIOPTERYGIDAE,TIPULIDAE)))

n.emerge_lake=with(l.fam.abund,(cbind(AEOLOSOMATIDAE,AMPULLARIIDAE,ANCYLIDAE,ARGULIDAE,ARRENURIDAE,ARTEMIIDAE,ASELLIDAE,ASTACIDAE,ATURIDAE,BALANIDAE,BELOSTOMATIDAE,BITHYNIIDAE,CAMBARIDAE,CHRYSOMELIDAE,CORBICULIDAE,COROPHIIDAE,CRANGONYCTIDAE,DREISSENIDAE,DRYOPIDAE,DYTISCIDAE,ELMIDAE,ENCHYTRAEIDAE,ERPOBDELLIDAE,EYLAIDAE,FELTRIIDAE,GAMMARIDAE,GLOSSIPHONIIDAE,GRAPSIDAE,GYRINIDAE,HALIPLIDAE,HELOPHORIDAE,HIRUDINIDAE,HYALELLIDAE,HYDRACHNIDAE,HYDRAENIDAE,HYDRIDAE,HYDROBIIDAE,HYDROCHIDAE,HYDRODROMIDAE,HYDROPHILIDAE,HYDROSCAPHIDAE,HYDRYPHANTIDAE,HYGROBATIDAE,KRENDOWSKIIDAE,LEBERTIIDAE,LEPTOCHELIIDAE,LIMNESIIDAE,LIMNOCHARIDAE,LUMBRICULIDAE,LYMNAEIDAE,LYNCEIDAE,MIDEIDAE,MIDEOPSIDAE,MOMONIIDAE,MYSIDAE,NAIDIDAE,NAUCORIDAE,NEPIDAE,NEREIDIDAE,NOTERIDAE,OEDICEROTIDAE,OXIDAE,PALAEMONIDAE,PENAEIDAE,PHYSIDAE,PIONIDAE,PISCICOLIDAE,PISIDIIDAE,PLANARIIDAE,PLANORBIDAE,PLEIDAE,PLEUROCERIDAE,PSEPHENIDAE,SABELLIDAE,SCIRTIDAE,SERPULIDAE,SPERCHONIDAE,SPERCHONTIDAE,SPHAEROMATIDAE,SPIONIDAE,TETRASTEMMATIDAE,THIARIDAE,THYASIDAE,TORRENTICOLIDAE,TUBIFICIDAE,UNIONICOLIDAE,UNIONIDAE,VALVATIDAE,VIVIPARIDAE,WETTINIDAE)))

l.tot.emerge=rowSums(emerge_lake) 
length(l.tot.emerge) #1210

l.emerge=l.tot.emerge/l.tot.abund
summary(l.emerge)

l.non.emerge <- rowSums(n.emerge_lake)
n.l.emerge=l.non.emerge/l.tot.abund
plot(n.l.emerge~l.emerge)


```

Binomial regresssion for emergence trait (LAKES)
```{r}
l.emerge_rows=which(dat4$FAMILY%in%colnames(emerge_lake))

l.emerge_binom=ifelse(dat4$FAMILY%in%colnames(emerge_lake),0,1)
dat4[999,]
l.emerge_binom[999]

names(dat4)
which(is.na(l.emerge_binom))
summary(l.emerge_binom)
summary(dat4$PTV)
?is.na
nrow(dat4)

indicator_emerge <- data.frame(l.emerge_binom,dat4$PTV)
colnames(indicator_emerge) <- c("l.emerge_binom","PTV")

binom2=glm(l.emerge_binom~PTV,data=indicator_emerge[-which(is.na(indicator_emerge$PTV)),],family="binomial")
summary(binom2)
confint(binom2)
exp(0.702)
exp(0.692)
exp(0.712)

```


Aggregate lake benthic macroinvert metrics (e.g., EPT_PIND,TOLRPIND)
```{r}
library(qqplotr)
names(dat4)
agg_lake_metrics=aggregate(dat4[,346:512],by=list(dat4$STATE),FUN=mean,na.rm=T)
head(agg_lake_metrics)

colnames(agg_lake_metrics)=c("STATE",names(agg_lake_metrics[2:ncol(agg_lake_metrics)]))

join_lakes_metrics <- inner_join(agg_lake_metrics,agg_lakeEmerge)
names(join_lakes_metrics)

cor(join_lakes_metrics$l.emerge,join_lakes_metrics$EPT_PIND)
cor(join_lakes_metrics$l.emerge,join_lakes_metrics$TOLRPIND)

```


Prep data for model selection using 'MuMIn'
```{r}
names(l.sites)

names(l.sites)
lakes_wq <- with(l.sites,aggregate(cbind(BASINAREA_KM2,PCT_DEVELOPED_BSN,PCT_AGRIC_BSN,SECMEAN,PH_LAB,COND,ANC,NTL,PTL,CL_PPM,SO4_PPM,CHLA),by=list(STATE),FUN=mean,na.rm=T)) # vars that correspond to common lake water quality metrics (e.g., trophic state metrics, conductivity, pH)
head(lakes_wq)

library(reshape2)
#install.packages('qqplotr')
library(qqplotr)

lmelt <- melt(lakes_norm,id=c(1:12),measure=2:ncol(lakes_wq))
lmelt <- melt(lakes_wq)
names(lmelt)

l_norm <- ggplot(data=lmelt,aes(sample=value)) + stat_qq_line() + stat_qq_point() + facet_wrap(~ variable,scales="free") + labs(x = "Theoretical Quantiles", y = "Sample Quantiles")
l_norm

# diagnostic plots show PH_LAB, SECMEAN close to normal; all others right-skewed, need log-transformation

names(lakes_wq)
log_agg_lakes=data.frame(lakes_wq[,1],log10(lakes_wq[,2:13]))
names(log_agg_lakes)
colnames(log_agg_lakes)=c("STATE",names(log_agg_lakes[2:13]))
summary(log_agg_lakes)
names(log_agg_lakes)

# check for multicollinearity problems.

vifcor(log_agg_lakes[,c(2:13)],th=0.75)

agg_lakeEmerge=aggregate(cbind(l.emerge,n.l.emerge),by=list(l.sites$STATE),FUN=mean,na.rm=T)
head(agg_lakeEmerge)
colnames(agg_lakeEmerge)=c("STATE","l.emerge","n.l.emerge")

```


Model selection process
```{r}
#install.packages('usdm')
library(usdm)

library('MuMIn')
names(agg_lakeEmerge)

agg_lakeEmerge=data.frame(agg_lakeEmerge,log_agg_lakes)
names(agg_lakeEmerge)

# emergent
glob.mod1=lm(l.emerge~BASINAREA_KM2+PCT_DEVELOPED_BSN+PCT_AGRIC_BSN+SECMEAN+PH_LAB+COND+ANC+NTL+PTL+CL_PPM+SO4_PPM+CHLA,data=agg_lakeEmerge)
dd.lake1=dredge(glob.mod1,beta="sd")
dd.lake.top1=subset(dd.lake1,delta<2)
dd.lake.avg1=model.avg(dd.lake.top1)
summary(dd.lake.avg1)

# non-emergent
glob.mod2=lm(n.l.emerge~BASINAREA_KM2+PCT_DEVELOPED_BSN+PCT_AGRIC_BSN+SECMEAN+PH_LAB+COND+ANC+NTL+PTL+CL_PPM+SO4_PPM+CHLA,data=agg_lakeEmerge)
dd.lake2=dredge(glob.mod2,beta="sd")
dd.lake.top2=subset(dd.lake2,delta<2)
dd.lake.avg2=model.avg(dd.lake.top2)
summary(dd.lake.avg2)

l.emerge_int=lm(agg_lakeEmerge$l.emerge~1)
summary(l.emerge_int)
AICc(l.emerge_int) # -59.209



```


Code chunks to produce lake emergent insect vs. lake water quality plots (plots not included in publication).
```{r}
names(agg_lakeEmerge)
chl <- ggplot() + geom_point(data=agg_lakeEmerge,aes(x=10^CHLA,y=l.emerge)) + geom_smooth(data=agg_lakeEmerge,aes(x=10^CHLA,y=l.emerge)) + theme_classic() + labs(x="Chlorophyll a",y="Rel. abund. emergent")
chl

ptl <- ggplot() + geom_point(data=agg_lakeEmerge,aes(x=10^PTL,y=l.emerge)) + geom_smooth(data=agg_lakeEmerge,aes(x=10^PTL,y=l.emerge)) + theme_classic() + labs(x="TP (mg/L)",y="") + scale_x_log10()
ptl

cl <- ggplot() + geom_point(data=agg_lakeEmerge,aes(x=10^CL_PPM,y=l.emerge)) + geom_smooth(data=agg_lakeEmerge,aes(x=10^CL_PPM,y=l.emerge)) + theme_classic() + labs(x="Chloride (mg/L)",y="") + scale_x_log10()
cl

```

