---
title: "bbs_annual_indices"
output: html_document
editor_options: 
  chunk_output_type: console
---


# USGS Breeding Bird Survey Analysis for Manning & Sullivan 2021
```{r load dplyr}

library(dplyr)

```

## USGS analyzed data (annual indices of abundance) can be found here: https://www.mbr-pwrc.usgs.gov/
```{r  load bbs data}
bbs_indices=read.csv("/Users/davidmanning/Documents/BBS_Annual_Indices_Estimates_2015_7-29-2016.csv")
head(bbs_indices) # check
str(bbs_indices) # check structure

```

## The following code chunk removes data from Canada, subsets BBS data to to aerial insectivore species, plus Mourning Dove, Goldfinch, Gray Catbird, White-breasted Nuthatch
```{r}

levels(bbs_indices$Region.Code) # check Region.Code
levels(bbs_indices$AOU.Number) # check AOU.Number
us_regions=c("ALA","ARI","ARK","CAL","COL","CON","DEL","FLA","GA ","IA ","IDA","ILL","IND","KAN","KY ","LA ","MAS","MD ","ME ","MIC","MIN","MIS","MO ","MON","NC ","ND ","NEB","NEV","NH ","NJ ","NM ","NY ","OHI","OKL","ORE","PA ","RI ","SC ","SD ","TEN","TEX","UT ","VA ","VT ","WAS","WIS","WV ","WYO") # vector of Region.Codes
bbs_indices[5794,] # check

spp_list=c("s04470","s04620","s04660","s04630","s04590","s04641","s06110","s04570","s06140","s04560","s04171","s04610","s04520","s04670","s04650","s04661","s06130","s04230","s04200","s04440","s05290","s03160","s07270","s07040","s06170","s06150") # list of aerial insectivore spp. + 'ref' spp.
length(spp_list) # check length

spp_names=c("WEKI","WEWP","WIFL","YBFL","OSFL","PSFL","PUMA","SAPH","TRES","EAPH","EWPW","EAWP","GCFL","LEFL","ACFL","ALFL","BARS","CHSW","CONI","EAKI","AMGO","MODO","WBNU","GRCA","NRWS","VGSW") # vector of spp abbreviations.
length(spp_names) # check length

habitat <- c("Upland","Upland","Riparian Obligate","Riparian Facultative","Riparian Facultative","Riparian Facultative","Riparian Facultative","Upland","Riparian Obligate","Riparian Facultative","Upland","Upland","Riparian Obligate","Upland","Riparian Obligate","Riparian Obligate","Riparian Facultative","Riparian Facultative","Riparian Facultative","Riparian Facultative","Non-aerial insectivore","Non-aerial insectivore","Non-aerial insectivore","Non-aerial insectivore","Non-aerial insectivore","Upland") # vector of habitats for bird species analyzed.

spp <- data.frame(spp_list,spp_names,habitat) # data.frame with AOU.Numbers lined up with spp abbreviations.
colnames(spp) <- c("AOU.Number","name","habitat") # change column names

usgs_spp=filter(bbs_indices,AOU.Number%in%spp_list) # subset the data to AOU.Numbers incl. in spp list.
usgs_us=filter(usgs_spp,Region.Code%in%us_regions) # subset to US states only.
sort(unique(usgs_us$Region.Code)) # check Region.Code

bbs_ref=usgs_us
names(usgs_us)

```

## code chunk subsets BBS dataset by year(s) that coincided with US EPA's Wadeable Streams Assessment (2000-2004), and National Lakes Assessment (2007) (plus 2008 and 2009)
```{r subset bbs dataset by year}

# note: PSFL not available in these data, likely because of discussion surrounding its sympatry with Cordilleran Flycatcher.

library(dplyr)

bbs_2000 <- bbs_ref %>%
  filter(Year=="2000") %>%
  group_by(AOU.Number,Region.Code) %>%
    summarise(ann_ind=mean(Annual.Index))

bbs_2001 <- bbs_ref %>%
  filter(Year=="2001") %>%
  group_by(AOU.Number,Region.Code) %>%
    summarise(ann_ind=mean(Annual.Index))

bbs_2002 <- bbs_ref %>%
  filter(Year=="2002") %>%
  group_by(AOU.Number,Region.Code) %>%
    summarise(ann_ind=mean(Annual.Index))

bbs_2003 <- bbs_ref %>%
  filter(Year=="2003") %>%
  group_by(AOU.Number,Region.Code) %>%
    summarise(ann_ind=mean(Annual.Index))

bbs_2004 <- bbs_ref %>%
  filter(Year=="2004") %>%
  group_by(AOU.Number,Region.Code) %>%
    summarise(ann_ind=mean(Annual.Index))
levels(bbs_2004$AOU.Number)

bbs_2005 <- bbs_ref %>%
  filter(Year=="2005") %>%
  group_by(AOU.Number,Region.Code) %>%
    summarise(ann_ind=mean(Annual.Index))

bbs_2006 <- bbs_ref %>%
  filter(Year=="2006") %>%
  group_by(AOU.Number,Region.Code) %>%
    summarise(ann_ind=mean(Annual.Index))

bbs_2007 <- bbs_ref %>%
  filter(Year=="2007") %>%
  group_by(AOU.Number,Region.Code) %>%
    summarise(ann_ind=mean(Annual.Index))

bbs_2008 <- bbs_ref %>%
  filter(Year=="2008") %>%
  group_by(AOU.Number,Region.Code) %>%
    summarise(ann_ind=mean(Annual.Index))

bbs_2009 <- bbs_ref %>%
  filter(Year=="2009") %>%
  group_by(AOU.Number,Region.Code) %>%
    summarise(ann_ind=mean(Annual.Index))
head(bbs_2009) # check

# combine all years
bbs_00_09 <- bbs_ref %>%
  filter(Year=="2000"|Year=="2009") %>%
  group_by(AOU.Number,Region.Code) %>%
    summarise(ann_ind=mean(Annual.Index))
nrow(bbs_00_09)

bbs_00_09 <- rbind(bbs_2000,bbs_2001,bbs_2002,bbs_2003,bbs_2004,bbs_2005,bbs_2006,bbs_2007,bbs_2008,bbs_2009)
nrow(bbs_00_09) # check rows

bbs_00_09$year <- c(rep("2000",837),rep("2001",837),rep("2002",837),rep("2003",837),rep("2004",837),rep("2005",837),rep("2006",837),rep("2007",837),rep("2008",837),rep("2009",837))



```

## compare annual index values among years
```{r}

bbs_year_lm <- lm(ann_ind~year,data=bbs_00_09)
summary(bbs_year_lm) # all p > 0.05

bbs_aov <- aov(ann_ind~year,data=bbs_00_09)
summary(bbs_aov)



```

## BBS Annual index values not different among years in this time span, proceed with analysis focusing on final year of WSA (2004) and NLA (2007). This chunk subsets the data to those years only and joins with aquatic insect data from EPA surveys.

``` {r}

bbs_2004_emerge <- inner_join(bbs_2004,agg_str_metrics) # join with aquatic insect data frames from 'EPA_emergence_2021_01_27_FrontEcolEvol.rmd'
unique(bbs_2004_emerge$AOU.Number) # check AOU.Number
bbs_2004_1 <- inner_join(bbs_2004_emerge,spp) # join above data frame to add species abbreviations
which(bbs_2004_1$ann_ind==0) # check for zeros
bbs_2004_1[94,] #inspect row with zero value
length(unique(bbs_2004_1$name)) #check

bbs_2007_emerge <- inner_join(bbs_2007,agg_p.emerge)
bbs_2007_1 <- inner_join(bbs_2007_emerge,spp)
which(bbs_2007_1$ann_ind==0) # check for zeros
bbs_2004_1[94,] #inspect row with zero value
length(unique(bbs_2004_1$name)) #check

```

``` {r check BBS data for normal distribution}

# q-q plots
qq_04 <- ggplot(data=bbs_2004_1,aes(sample=ann_ind)) + stat_qq_line() + stat_qq_point() + labs(x="Theoretical Quantiles",y="Sample Quantiles",title="BBS 2004 Q-Q plot")
qq_04

qq_07 <- ggplot(data=bbs_2007_1,aes(sample=ann_ind)) + stat_qq_line() + stat_qq_point() + labs(x="Theoretical Quantiles",y="Sample Quantiles",title="BBS 2007 Q-Q plot")
qq_07

# Kolmogorov-Smirnov tests
ks.test(bbs_2004_1$ann_ind[-94],pnorm) # D = 0.545, P << 0.001
ks.test(bbs_2007_1$ann_ind[-94],pnorm) # D = 0.542, P << 0.001


```


```{r}
par(mfrow=c(1,1))
boxplot(Annual.Index~AOU.Number,data=bbs_ref)
summary(bbs_ref)

zero_ind=which(bbs_ref$Annual.Index==0)
bbs_rem=bbs_ref[-zero_ind,]

agg_p.emerge$Region.Code <- c("ALA","ARI","ARK","CAL","COL","CON","DEL","FLA","GA ","IDA","ILL","IND","IA ","KAN","KY ","LA ","ME ","MD ","MAS","MIC","MIN","MIS","MO ","MON","NEB","NEV","NH ","NJ ","NM ","NY ","NC ","ND ","OHI","OKL","ORE","PA ","RI ","SC ","SD ","TEN","TEX","UT ","VT ","VA ","WAS","WV ","WIS","WYO")

agg_str_metrics$Region.Code <- c("ALA","ARI","ARK","CAL","COL","CON","DEL","FLA","GA ","IDA","ILL","IND","IA ","KAN","KY ","LA ","ME ","MD ","MAS","MIC","MIN","MIS","MO ","MON","NEB","NEV","NH ","NJ ","NM ","NY ","NC ","ND ","OHI","OKL","ORE","PA ","RI ","SC ","SD ","TEN","TEX","UT ","VT ","VA ","WAS","WV ","WIS","WYO")


```

## Bayesian models testing for relationships between bird species annual indices of abundance and relative abundance of emergent insects (streams + lakes)
```{r}

library(brms)

# species-level model (random slope for each aerial insectivore species)
brm04 <- brm(log10(ann_ind) ~ p.emerge + (1 + p.emerge|name),data=bbs_2004_1[-94,],family=gaussian,chains=4,iter=2000,warmup=1000)
fixef(brm04,probs=c(0.05,0.95)) # check fixed effects
ranef(brm04,probs=c(0.05,0.95)) # check random effects
str04_coef <- data.frame(coef(brm04,probs=c(0.05,0.95))) # create data.frame of spp-level coefficients

# habitat-level model (random slope for each habitat guild: riparian obligate, riparian facultative, upland, non-aerial insectivore)
brm04_hab <- brm(log10(ann_ind) ~ p.emerge + (1 + p.emerge|habitat) ,data=bbs_2004_1[-94,],family=gaussian,chains=4,iter=2000,warmup=1000)
fixef(brm04_hab)
coef(brm04_hab,probs=c(0.05,0.95))

```

## Bayesian models for relationships between BBS annual indices of abundance and Ephemeroptera, Plecoptera, Trichoptera (EPA metric = 'EPT_PIND'), and percent tolerant individuals (EPA metric = 'TOLRPIND') found in streams.
```{r}
bbs_2004_1$pEPT=bbs_2004_1$EPT_PIND/100 # convert from % to proportion
bbs_2004_1$pTOL <- bbs_2004_1$TOLRPIND/100 # convert from % to proportion

brm04_ept <- brm(log10(ann_ind) ~ pEPT + (1 + pEPT|name),data=bbs_2004_1[-94,],family=gaussian,chains=4,iter=2000,warmup=1000)
coef(brm04_ept,probs=c(0.05,0.95))
str04_ept_coef <- data.frame(coef(brm04_ept,probs=c(0.05,0.95)))

brm04_tol <- brm(log10(ann_ind) ~ pTOL + (1 + pTOL|name),data=bbs_2004_1[-94,],family=gaussian,chains=4,iter=2000,warmup=1000)
fixef(brm04_tol)
coef(brm04_tol,probs=c(0.05,0.95))

str04_tol_coef <- data.frame(coef(brm04_tol,probs=c(0.05,0.95)))

str04_coef <- data.frame(coef(brm04,probs=c(0.05,0.95)))
str04_ept_coef <- data.frame(coef(brm04_ept,probs=c(0.05,0.95)))
str04_tol_coef <- data.frame(coef(brm04_tol,probs=c(0.05,0.95)))


```


### Code to produce Figure 3 (coefficients of bird responses to stream insects by species)
```{r}

names(str04_coef)

str04_coef$spp_abb <- c("ACFL","ALFL","AMGO","BARS","CHSW","CONI","EAKI","EAPH","EAWP","EWPW","GCFL","GRCA","LEFL","MODO","NRWS","OSFL","PUMA","SAPH","TRES","VGSW","WBNU","WEKI","WEWP","WIFL","YBFL") # add spp abbreviations to the data.frame

str04_coef$Habitat <- c("RO","RO","NON","RF","RF","RF","RF","RF","UP","UP","RO","NON","UP","NON","RO","RF","RF","UP","RO","UP","NON","UP","UP","RO","RF") # add habitat to the data.frame

colnames(str04_coef) <- c("Int","ErrInt","Q5Int","Q95Int","Est","Err","Q5","Q95","spp_abb","Habitat") # rename columns
head(str04_coef) # check

order_ests <- str04_coef %>%
  arrange(Est,spp_abb) %>%
  mutate(name=factor(spp_abb, levels=spp_abb)) # sorts the slopes in numerical order

stp <- ggplot() + geom_point(data=order_ests,aes(x=Est,y=name,color=Habitat),size=3) + theme_classic() + geom_segment(data=order_ests,aes(x=Q5,y=name,xend=Q95,yend=name,color=Habitat)) + geom_vline(xintercept=0,col="gray") + labs(y="",x="Estimate") + theme(text=element_text(size=20)) + theme(legend.position="none")
stp

# add species abbreviations and habitats to stream EPT data.frame.
str04_ept_coef$spp_abb <- c("ACFL","ALFL","AMGO","BARS","CHSW","CONI","EAKI","EAPH","EAWP","EWPW","GCFL","GRCA","LEFL","MODO","NRWS","OSFL","PUMA","SAPH","TRES","VGSW","WBNU","WEKI","WEWP","WIFL","YBFL")
str04_ept_coef$Habitat <- c("RO","RO","NON","RF","RF","RF","RF","RF","UP","UP","RO","NON","UP","NON","RO","RF","RF","UP","RO","UP","NON","UP","UP","RO","RF")
colnames(str04_ept_coef) <- c("Int","ErrInt","Q5Int","Q95Int","Est","Err","Q5","Q95","spp_abb","Habitat")
head(str04_ept_coef)

order_ept_ests <- str04_ept_coef %>%
  arrange(Est,spp_abb) %>%
  mutate(name=factor(spp_abb, levels=spp_abb)) # sort slopes in numerical order.

st_ept <- ggplot() + geom_point(data=order_ept_ests,aes(x=Est,y=name,color=Habitat),size=3) + theme_classic() + geom_segment(data=order_ept_ests,aes(x=Q5,y=name,xend=Q95,yend=name,col=Habitat)) + geom_vline(xintercept=0,col="gray") + labs(y="",x="") + theme(text=element_text(size=20)) + theme(legend.position="none")
st_ept

# add species abbreviations and habitats to stream TOLRPIND data.frame.
str04_tol_coef$spp_abb <- c("ACFL","ALFL","AMGO","BARS","CHSW","CONI","EAKI","EAPH","EAWP","EWPW","GCFL","GRCA","LEFL","MODO","NRWS","OSFL","PUMA","SAPH","TRES","VGSW","WBNU","WEKI","WEWP","WIFL","YBFL")
str04_tol_coef$Habitat <- c("RO","RO","NON","RF","RF","RF","RF","RF","UP","UP","RO","NON","UP","NON","RO","RF","RF","UP","RO","UP","NON","UP","UP","RO","RF")
colnames(str04_tol_coef) <- c("Int","ErrInt","Q5Int","Q95Int","Est","Err","Q5","Q95","spp_abb","Habitat")
str04_tol_coef[,"Est"] # check

order_tol_ests <- str04_tol_coef %>%
  arrange(Est,spp_abb) %>%
  mutate(name=factor(spp_abb, levels=spp_abb)) # sort slopes in numerical order

st_tol <- ggplot() + geom_point(data=order_tol_ests,aes(x=Est,y=name,color=Habitat),size=3) + theme_classic() + geom_segment(data=order_tol_ests,aes(x=Q5,y=name,xend=Q95,yend=name,col=Habitat)) + geom_vline(xintercept=0,col="gray") + labs(y="",x="Coefficient") + theme(text=element_text(size=20)) + theme(legend.position="right")
st_tol


### put figures in multi-panel plot.
stp <- ggplot() + geom_point(data=order_ests,aes(x=Est,y=name,color=Habitat),size=3) + theme_classic() + geom_segment(data=order_ests,aes(x=Q5,y=name,xend=Q95,yend=name,color=Habitat)) + geom_vline(xintercept=0,col="gray") + labs(y="",x="Coefficient",title="") + theme(text=element_text(size=20)) + theme(legend.position="none") + annotate("text", label = "Emergent", x = 2.5, y = 1.5, size = 5, colour = "black")
stp

st_ept <- ggplot() + geom_point(data=order_ept_ests,aes(x=Est,y=name,color=Habitat),size=3) + theme_classic() + geom_segment(data=order_ept_ests,aes(x=Q5,y=name,xend=Q95,yend=name,col=Habitat)) + geom_vline(xintercept=0,col="gray") + labs(y="",x="Coefficient",title="") + theme(text=element_text(size=20)) + theme(legend.position="none") + annotate("text", label = "EPT", x = 2.5, y = 1.5, size = 5, colour = "black")
st_ept

st_tol <- ggplot() + geom_point(data=order_tol_ests,aes(x=Est,y=name,color=Habitat),size=3) + theme_classic() + geom_segment(data=order_tol_ests,aes(x=Q5,y=name,xend=Q95,yend=name,col=Habitat)) + geom_vline(xintercept=0,col="gray") + labs(y="",x="Coefficient",title="") + theme(text=element_text(size=20)) + theme(legend.position="right") +  annotate("text", label = "Tolerant", x = 2.5, y = 1.5, size = 5, colour = "black")
st_tol


ggarrange(stp,st_ept,st_tol,labels=c("a","b","c"),nrow=1)


```



## BBS annual indices of abundance from 2007 vs. aquatic insect data from EPA National Lakes Assessment survey.
```{r}

agg_lakeEmerge # data.frame from 'NLA_MACROS_2021_01_27_FrontEcolEvol.rmd'
agg_lakeEmerge$Region.Code <- c("ALA","ARI","ARK","CAL","COL","CON","DEL","FLA","GA ","IDA","ILL","IND","IA ","KAN","KY ","LA ","ME ","MD ","MAS","MIC","MIN","MIS","MO ","MON","NEB","NEV","NH ","NJ ","NM ","NY ","NC ","ND ","OHI","OKL","ORE","PA ","RI ","SC ","SD ","TEN","TEX","UT ","VT ","VA ","WAS","WV ","WIS","WYO") # add state abbreviations to data.frame

bbs_2007_lake <- inner_join(bbs_2007,join_lakes_metrics)
bbs_2007_l <- inner_join(bbs_2007_lake,spp)
which(bbs_2007_1$ann_ind==0)

l07<- ggplot() + geom_point(data=bbs_2007_l[-94,],aes(x=l.emerge,y=ann_ind)) + geom_smooth(data=bbs_2007_l[-94,],aes(x=l.emerge,y=ann_ind),method="lm") + facet_wrap(~name) + theme_bw() + xlab("Rel. abund. emergent") + ylab("State-level mean annual index") + scale_y_log10()
l07

# Bayesian model relating BBS annual indices to lake emergent insect values.
brm_l07 <- brm(log10(ann_ind) ~ l.emerge + (1 + l.emerge|name),data=bbs_2007_l[-94,],family=gaussian,chains=4,iter=2000,warmup=1000)
coef(brm_l07,probs=c(0.05,0.95))
?brm

lakes07_coef <- data.frame(coef(brm_l07,probs=c(0.05,0.95))) # data.frame of species-level slopes.

names(bbs_2007_l)

# convert lake 'EPT_PIND' and 'TOLRPIND' from % to proportion.
bbs_2007_l$pEPT <- bbs_2007_l$EPT_PIND/100
bbs_2007_l$pTOL <- bbs_2007_l$TOLRPIND/100

brm_l07_ept <- brm(log10(ann_ind) ~ pEPT + (1 + pEPT|name),data=bbs_2007_l[-94,],family=gaussian,chains=4,iter=2000,warmup=1000)
coef(brm_l07_ept,probs=c(0.05,0.95))

lakes07_ept_coef <- data.frame(coef(brm_l07_ept,probs=c(0.05,0.95)))

brm_l07_tol <- brm(log10(ann_ind) ~ pTOL + (1 + pTOL|name),data=bbs_2007_l[-94,],family=gaussian,chains=4,iter=2000,warmup=1000,control=list(adapt_delta=0.95))
coef(brm_l07_tol,probs=c(0.05,0.95))

lakes07_tol_coef <- data.frame(coef(brm_l07_tol,probs=c(0.05,0.95)))

# habitat-level model for lakes.
brm_l07_hab <- brm(log10(ann_ind) ~ l.emerge + (1 + l.emerge|habitat), data=bbs_2007_l[-94,],family=gaussian,chains=4,iter=2000,warmup=1000)
fixef(brm_l07_hab,probs=c(0.05,0.95))
coef(brm_l07_hab,probs=c(0.05,0.95))


```


## Code chunk to produce Figure 4 (coefficients of species-level bird responses to aquatic insects from lakes)
```{r}

#bbs_coefs <- rbind(str04_coef,str04_ept_coef,str04_tol_coef,lake07_coef,lake07_ept_coef,lake07_tol_coef)
#head(bbs_coefs)
names(lakes07_coef)
lakes07_coef


lakes07_coef$spp_abb <- c("ACFL","ALFL","AMGO","BARS","CHSW","CONI","EAKI","EAPH","EAWP","EWPW","GCFL","GRCA","LEFL","MODO","NRWS","OSFL","PUMA","SAPH","TRES","VGSW","WBNU","WEKI","WEWP","WIFL","YBFL") # add common name abbreviations
lakes07_coef$Habitat <- c("RO","RO","NON","RF","RF","RF","RF","RF","UP","UP","RO","NON","UP","NON","RO","RF","RF","UP","RO","UP","NON","UP","UP","RO","RF") # add habitat category
head(lakes07_coef)
colnames(lakes07_coef) <- c("Int","ErrInt","Q5Int","Q95Int","Est","Err","Q5","Q95","spp_abb","Habitat") # change colnames
head(lakes07_coef) # check

lakes_order_ests <- lakes07_coef %>%
  arrange(Est,spp_abb) %>%
  mutate(name=factor(spp_abb, levels=spp_abb)) # sort slopes in numerical order


lkp <- ggplot() + geom_point(data=lakes_order_ests,aes(x=Est,y=name,color=Habitat),size=3) + theme_classic() + geom_segment(data=lakes_order_ests,aes(x=Q5,y=name,xend=Q95,yend=name,color=Habitat)) + geom_vline(xintercept=0,col="gray") + labs(y="",x="Estimate") + theme(text=element_text(size=20)) + theme(legend.position="none")
lkp

# lakes EPT
lakes07_ept_coef$spp_abb <- c("ACFL","ALFL","AMGO","BARS","CHSW","CONI","EAKI","EAPH","EAWP","EWPW","GCFL","GRCA","LEFL","MODO","NRWS","OSFL","PUMA","SAPH","TRES","VGSW","WBNU","WEKI","WEWP","WIFL","YBFL")
lakes07_ept_coef$Habitat <- c("RO","RO","NON","RF","RF","RF","RF","RF","UP","UP","RO","NON","UP","NON","RO","RF","RF","UP","RO","UP","NON","UP","UP","RO","RF")
colnames(lakes07_ept_coef) <- c("Int","ErrInt","Q5Int","Q95Int","Est","Err","Q5","Q95","spp_abb","Habitat")
head(lakes07_ept_coef)

lakes_order_ept_ests <- lakes07_ept_coef %>%
  arrange(Est,spp_abb) %>%
  mutate(name=factor(spp_abb, levels=spp_abb)) # sort slopes in numerical order

lake_ept <- ggplot() + geom_point(data=lakes_order_ept_ests,aes(x=Est,y=name,color=Habitat),size=3) + theme_classic() + geom_segment(data=lakes_order_ept_ests,aes(x=Q5,y=name,xend=Q95,yend=name,col=Habitat)) + geom_vline(xintercept=0,col="gray") + labs(y="",x="") + theme(text=element_text(size=20)) + theme(legend.position="none")
lake_ept

# lakes 'TOLRPIND'
lakes07_tol_coef$spp_abb <- c("ACFL","ALFL","AMGO","BARS","CHSW","CONI","EAKI","EAPH","EAWP","EWPW","GCFL","GRCA","LEFL","MODO","NRWS","OSFL","PUMA","SAPH","TRES","VGSW","WBNU","WEKI","WEWP","WIFL","YBFL")
lakes07_tol_coef$Habitat <- c("RO","RO","NON","RF","RF","RF","RF","RF","UP","UP","RO","NON","UP","NON","RO","RF","RF","UP","RO","UP","NON","UP","UP","RO","RF")
colnames(lakes07_tol_coef) <- c("Int","ErrInt","Q5Int","Q95Int","Est","Err","Q5","Q95","spp_abb","Habitat")
lakes07_tol_coef[,"Est"] # check

lakes_order_tol_ests <- lakes07_tol_coef %>%
  arrange(Est,spp_abb) %>%
  mutate(name=factor(spp_abb, levels=spp_abb)) # sort slopes in numerical order

lakes_tol <- ggplot() + geom_point(data=lakes_order_tol_ests,aes(x=Est,y=name,color=Habitat),size=3) + theme_classic() + geom_segment(data=lakes_order_tol_ests,aes(x=Q5,y=name,xend=Q95,yend=name,col=Habitat)) + geom_vline(xintercept=0,col="gray") + labs(y="",x="Coefficient") + theme(text=element_text(size=20)) + theme(legend.position="right")
lakes_tol


### combine into multi-panel figure.
lkp <- ggplot() + geom_point(data=lakes_order_ests,aes(x=Est,y=name,color=Habitat),size=3) + theme_classic() + geom_segment(data=lakes_order_ests,aes(x=Q5,y=name,xend=Q95,yend=name,color=Habitat)) + geom_vline(xintercept=0,col="gray") + labs(y="",x="Coefficient") + theme(text=element_text(size=20)) + theme(legend.position="none") + annotate("text", label = "Emergent", x = 2.5, y = 1.5, size = 5, colour = "black")
lkp

lakes_ept <- ggplot() + geom_point(data=lakes_order_ept_ests,aes(x=Est,y=name,color=Habitat),size=3) + theme_classic() + geom_segment(data=lakes_order_ept_ests,aes(x=Q5,y=name,xend=Q95,yend=name,col=Habitat)) + geom_vline(xintercept=0,col="gray") + labs(y="",x="Coefficient") + theme(text=element_text(size=20)) + theme(legend.position="none") +  annotate("text", label = "EPT", x = 7, y = 1.5, size = 5, colour = "black")
lakes_ept

lakes_tol <- ggplot() + geom_point(data=lakes_order_tol_ests,aes(x=Est,y=name,color=Habitat),size=3) + theme_classic() + geom_segment(data=lakes_order_tol_ests,aes(x=Q5,y=name,xend=Q95,yend=name,col=Habitat)) + geom_vline(xintercept=0,col="gray") + labs(y="",x="Coefficient") + theme(text=element_text(size=20)) + theme(legend.position="right") + annotate("text", label = "Tolerant", x = 2, y = 1.5, size = 5, colour = "black")
lakes_tol

ggarrange(lkp,lakes_ept,lakes_tol,labels=c("a","b","c"),nrow=1)

```



